---
title: "ESM262 Notes"
author: "Ignacia Rivera"
date: "May 5, 2017"
output: html_notebook
---

## Tidying data

### Gather and Spread

When we don't have “a variable per column, one observation per row” we can use gather and spread.The function requires the columns you want to gather using ``, a Key, which is the new variable we want to create and repeat and the value which is the second variable paired to the key. 

Spread take the values in a column an transform them into a column titles. 

```{r cars}

library(tidyverse)

View(table4a)

tidy4a <- table4a %>% 
  gather(`1999`, `2000`, key = "year", value = "cases")

View(tidy4a)
```

### Separate

When we do not have a value per cell we use separate. 

```{r pressure, echo=FALSE}
View(table3)

tidy3<- table3 %>% 
  separate(rate, into = c("cases", "population"), sep = "/")

View(tidy3)
```

### Unite 

```{r}
View(table5)

tidy5 <- table5 %>% 
  unite(new, century, year, sep = "")

```

## Missing values

Here is a data frame with implicit (first quarter of 2016) and explicit missing values (NA). Is always good to have one of the two forms of missing data. complete do combinations of all the possible values that exist for a variable and fill the value of the combinations with no values with an NA. 

```{r}

stocks <- tibble(
  year   = c(2015, 2015, 2015, 2015, 2016, 2016, 2016),
  qtr    = c(   1,    2,    3,    4,    2,    3,    4),
  return = c(1.88, 0.59, 0.35,   NA, 0.92, 0.17, 2.66)

  )

View(stocks)

# Option 1- Explicit representation of mising values

stocks1 <- stocks %>% 
  spread(year, return)

View(stocks1)

# Option 2- Implicit representation of missing values 

stocks2 <- stocks %>% 
  spread(year, return) %>% 
  gather(year, return, `2015`:`2016`, na.rm = TRUE)

View (stocks2)

#Option 3- Explicit representation of missing values

stocks3 <- stocks %>%
  complete(year,qtr)

View(stocks3)

```


## Transforming data

Once ww ave our data "tidied" enough we can start working with it. For doing this, we usually use dplyr basic functions like: 

Pick observations by their values (filter()).
Reorder the rows (arrange()).
Pick variables by their names (select()).
Create new variables with functions of existing variables (mutate()).
Collapse many values down to a single summary (summarise()).
Summarizing stuff for more than just one column we use (summarise_all()).

```{r}

library(nycflights13)

View(flights)

flights_firstdaymonth <- flights %>% 
  filter(month ==1, day == 1)

#or

flights_firstdaymonth <- flights %>% 
  filter(month ==1) %>%
  filter(day == 1)

View(flights_firstdaymonth)

dates <- flights %>% 
  select(yyyy = year, mm = month, dd= day)

View(dates)

delays <- flights %>%
  group_by(dest) %>%
  summarise(
    count = n(),
    dist= mean(distance, na.rm= TRUE), 
    delay= mean(arr_delay, na.rm= TRUE)
  ) %>%
  filter(count <20)

 

```

## Data visualization - ggplot

















