---
title: 'HW2: Data Ingest'
author: "Ignacia Rivera"
date: "May 9, 2017"
output:
  pdf_document: default
  html_document: default
---

```{r setup, message= FALSE}

library(tidyverse)
library(knitr)

```

# Importing and tidying data

```{r database, message = FALSE}

#Importing

gaz_raw <- read_delim("C:/Users/Ignacia Rivera/Dropbox/MESM/Quarter VI/ComputingForEnvScience/Rivera_ESM262/HW2/CA_Features_20170401.txt", delim= "|")

#Selecting variables of interest and formating date

gaz <- gaz_raw %>% 
  select (-COUNTY_NUMERIC, 
          -STATE_NUMERIC, 
          -PRIMARY_LAT_DMS, 
          -PRIM_LONG_DMS, 
          -SOURCE_LAT_DMS, 
          -SOURCE_LONG_DMS, 
          -ELEV_IN_FT) %>% 
  mutate(DATE_CREATED = as.Date(DATE_CREATED, format= "%m/%d/%Y"), 
         DATE_EDITED = as.Date(DATE_EDITED, format= "%m/%d/%Y" ))

#Assigning NA to missed values
#Is this is the fastest way of replece them ?

gaz$PRIM_LAT_DEC <- ifelse(gaz$PRIM_LAT_DEC == 0, NA, gaz$PRIM_LAT_DEC)
gaz$PRIM_LONG_DEC <- ifelse(gaz$PRIM_LONG_DEC == 0, NA, gaz$PRIM_LONG_DEC)
gaz$MAP_NAME <- ifelse(gaz$MAP_NAME =='Unknown', NA, gaz$MAP_NAME)

#Selecting observations of interest

gaz <- gaz %>%
  filter(!(PRIM_LAT_DEC == 'NA' | PRIM_LONG_DEC == 'NA')) %>% 
  filter(STATE_ALPHA == 'CA')

#Saving the clean tibble

write_delim(gaz, "C:/Users/Ignacia Rivera/Dropbox/MESM/Quarter VI/ComputingForEnvScience/Rivera_ESM262/HW2/gaz.csv", delim= "|")

```

## Analyzing data

```{r Analysis, results= "hide"}

#What is the most-frequently-occuring feature name?

count_feature_name <- count(gaz, FEATURE_NAME)
Ans1 <- count_feature_name[which.max(count_feature_name$n), 1]

#What is the least-frequently-occuring feature class?

count_feature_class <- count(gaz, FEATURE_CLASS)
Ans2 <- count_feature_class[which.min(count_feature_class$n), 1]

#What is the approximate center point of each county?

county_center <- gaz %>%
 group_by(COUNTY_NAME) %>%
  summarise(
    latMin = min(PRIM_LAT_DEC, na.rm= TRUE),
    latMax = max(PRIM_LAT_DEC, na.rm= TRUE),
    longMin = min(PRIM_LONG_DEC, na.rm= TRUE),
    longMax = max(PRIM_LONG_DEC, na.rm= TRUE)) %>% 
  mutate(lat_center = (latMin + latMax) /2)%>% 
  mutate(long_center = (longMin + longMax)/2) %>% 
  select(COUNTY_NAME, lat_center, long_center)

#What are the fractions of the total number of features in each county that are natural? man-made?

natural <- c("Valley", "Spring" ,"Stream", "Gut", "Canal", "Cape", "Lake", "Summit", "Plain", "Woods",   "Gap",   "Ridge","Bay",  "Beach",   "Channel", "Slope", "Island",  "Basin",   "Cliff", "Swamp", "Area" , "Bend",  "Range", "Falls",  "Forest", "Lava", "Rapids",   "Crater", "Glacier", "Oilfield","Sea", "Arroyo","Isthmus")

Feature_category <- tibble(FEATURE_CLASS= unique(gaz$FEATURE_CLASS)) %>% 
  mutate(category = ifelse(FEATURE_CLASS %in% natural == "TRUE", 1, 0)) #1 = natural, 0= man-made
  
gaz_complete <- inner_join(gaz, Feature_category, by = "FEATURE_CLASS")

Features_type <- gaz_complete%>% 
  group_by(COUNTY_NAME) %>% 
  summarise(total = n(), natural = sum(category)) %>% 
  mutate(fraction_natural = natural/total) %>% 
  mutate(fraction_manmade = 1 - fraction_natural) %>% 
  select(COUNTY_NAME, fraction_natural, fraction_manmade)

```

## What is the most-frequently-occuring feature name?

 The most-frequently- occuring feature in California is `r Ans1`.
 
## What is the least-frequently-occuring feature class?

 The least-frequently-occuring feature in California is `r Ans2`.
 
## What is the approximate center point of each county?

```{r table 1, echo = FALSE}

kable(county_center, caption = "Longitude and latitude of the center of each county in California.")
```

## What are the fractions of the total number of features in each county that are natural? man-made?

```{r table 2, echo = FALSE}

kable(Features_type, caption = "Fraction of natural and man-made features in each county in California.")
```
